---
title: "Analysis"
author: "Andrew Redd, PhD."
date: "R Bootcamp 2020"
output: 
    ioslides_presentation:
        logo: "images/U_Health_stacked_png_red.png"
        theme: "paper"
        css: "style.css"
        keep_md: FALSE
        md_extensions: 
            - "+task_lists"
            - "+multiline_tables"
        includes:
            in_header: timer.htm
        standalone: TRUE
        widescreen: TRUE
---

```{r setup, include=FALSE}
library(knitr)
library(printr)
library(tidyverse)
library(magrittr)
knitr::opts_chunk$set(echo = TRUE, error=TRUE
                     , class.error='error'
                     , class.warning='warning'
                     , class.message='message'
                     , cache=FALSE)
options( repos="https://cran.rstudio.org"
       , width=60
       )
```

# Hypothesis Tests

## T-test

### Something Ridiculous
```{r include=FALSE}
load(file="data/our.data.RData")
data <- 
    our.data %>%
        filter(SheetName == "Jan 06, 2016") %>% 
        mutate( in.Africa = ifelse( iso3c %in% c('ESP', 'GBR', 'USA', 'ITA')
                                  , 'Other', 'Africa')
              )
```

Test for the difference in GDP Per capita.
```{r}
# using the restriocted data from the exercise
(gdp.test <- t.test(NY.GDP.PCAP.CD~in.Africa, data=data))
```

-----
:::{.terminology}
Formulas, created with a `~`, represent relationships. They can 

* be one sided `~x`
    - Lambda functions `~log(.+1)`
* or two sided `y ~ x + z`
    - specify relationships or models
* often include functions
    - `y ~ x + log(z)`
    - `y ~ x + poly(z,3)`, a polynomial fit of degree 3 on `z`
* have special syntax
    - interaction `y ~ x:z`
    - crossing `y ~ a*b` is equal to `y ~ a + b + a:b`
    - nesting `y ~ a + b %in% a` or equivalently `y~a/b`
:::

## Getting usable results from a model {.build}

The `gdp.test` object is a `r class(gdp.test)` object, 
which prints nicely but what if we want to include this in our table 1?

Try these:
```r
str(gdp.test) # get the underlying structure of the object.
glimpse(gdp.test) # alternative to str that handles some objects better. 
gdp.test$p.value
getElement(gdp.test, 'p.value')
gdp.test[['p.value']]
```

## Exercise: Proportion Test

Use `prop.test` to test for the difference in proportions of males for all countries.

:::{#timer3 .timer onclick="startTimer()" warnafter=30}
5:00
:::


## Solution


```{r}
suppressMessages(library(magrittr))
data %>% 
    mutate(N.males = round(SP.POP.TOTL*SP.POP.TOTL.MA.ZS/100)) %$%
    prop.test(N.males, SP.POP.TOTL)
```

## Solution, what I did

```r
suppressMessages(library(magrittr)) #< To ignore unimportant message.
data %>% 
    mutate(N.males = round(SP.POP.TOTL*SP.POP.TOTL.MA.ZS/100)) %$%
    prop.test(N.males, SP.POP.TOTL)
```

0. Load magrittr to get the accessor pipe `%$%`.
1. `mutate(N.male = ...)`, get counts of males, because prop.test expects counts.
2. used `%$%` the Accessor pipe to make the variables directly accessible to the next command
3. used `prop.test` to perform the test.

## Alternate Version

Passing in a matrix with 'successes' and 'failures':
```{r}
data %>% 
    # get our 'successes' and 'failures'
    transmute( males = round(SP.POP.TOTL*SP.POP.TOTL.MA.ZS/100)
             , females = SP.POP.TOTL-males
             ) %>%
    # convert to a matrix
    as.matrix() %>% 
    prop.test()
```


# Linear Models

## Bordeaux Wine Data
```{r}
wine <- read.csv("data/Bordeaux.csv")
```

### Variables
* `Wine` - the name of the wine.
* `Price` - The price of the wine in pounds sterling(&pound;)
* `ParkerPoints` - the rating out of 100 given by [Robert Parker](https://www.robertparker.com/).
* `CoatesPoints` - the rating out of 20 given by [Clive Coates](http://www.clive-coates.com/)
* `P95andAbove` - a dummy variable, 1 if `ParkerPoints>=95`
* `FirstGrowth` - inditcator of if the wine is a [first growth](https://en.wikipedia.org/wiki/First_Growth)
* `CultWine` - indicator of if the wine is a [cult wine](https://en.wikipedia.org/wiki/Cult_wine)
* `Pomerol` - indicator for if the wine is from [Pomerol](https://en.wikipedia.org/wiki/Pomerol), France
* `VintageSuperstar` - indicator if the wine is a superstar


## Wine Model

We will rely on the normal approximation for proportions.

```{r}
model <- lm( Price ~ . - Wine - P95andAbove, data=wine)
model
```
Not really useful.

## Wine Model

We will rely on the normal approximation for proportions.

```r
model <- lm( Price ~ . - Wine - P95andAbove, data=wine)
model
```
::: {.concept}

Fomula Subtraction

: `Price ~ . - Wine - P95andAbove` should be read as

    > "Model Price by all variables **except** Wine and P95andAbove."

:::

## Summarizing Models

```{r}
(model.summary <- summary(model))
```

## Exercise: Try the following

:::::: {.columns-2}
:::{.column}
#### Extracting Parts of the model

```r
coef(model)
coef(summary(model))
deviance(model)
formula(model)
residuals(model)
```
:::
:::{.column}
#### Operations on models

```r
summary(model)
plot(model)
predict(model)
vcov(model)
anova(model)
aov(model)
```
:::
::::::

:::{#modelstimer .timer onclick="startTimer()" warnafter=30}
5:00
:::

## Residual plots

```{r h=4,w=6}
plot(model, 1) #< Residuals vs. Fitted  | Goodness of fit
```

## Residual plots

```{r}
plot(model, 2) #< Normal Q-Q            | Appropriate error model 
```

## Residual plots

```{r}
plot(model, 3) #< Scale-Location        | Homoscedasticity
```

## Residual plots

```{r}
plot(model, 5) #< Residual vs. Leverage | Influential points
```
