---
title: "Geospatial example"
author: "Andrew Redd"
date: "8/20/2020"
output: html_document
---

# Packages

You will have to install:

* `rnaturalearth`
* `rnaturalearthdata`
* `rgeos`

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(viridis)
library(rnaturalearth)
requireNamespace('rgeos')
```

# Data

Load our data with world bank features.

```{r}
load("data/our.data.RData")
```

Get world data geo-features from `rnatualworld::ne_countries`

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)
glimpse(world)
```
 Join the data together and convert to a `sf` which is geometric simle features data frame.

```{r}
plot.data <- inner_join(our.data, world, c('iso3c'='iso_a3')) %>% sf::st_sf()
```

# Plotting

We use `geom_sf` to create a geo-spatial layer.  Here we create a base layer.

```{r}
ggplot() + geom_sf(data=world)
```

Then add our data layer with fill set to `All.cases` and use the continuous viridis scale for visibility.

```{r}
ggplot() + geom_sf(data=world) + geom_sf(aes(fill=All.cases), data = plot.data) +
    scale_fill_viridis_c()
```
 Let's limit the view to just Africa and log transform the scale.
```{r}
ggplot() + 
    geom_sf(data = world %>% filter(continent == 'Africa')) +
    geom_sf(aes(fill=All.cases), data = plot.data %>% filter(continent == 'Africa')) +
    scale_fill_viridis_c(trans = 'log10')
```
